---
title: "Assignment 3"
author: "Olaleye Damilare"
date: "02/06/20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

#### Regression Analysis

Data is from the ```mpg``` dataset in the ```ggplot2``` package.  Interested in predicting the highway mileage (variable ```hwy```) from the engine displacement (variable ```displ```).

We start by constructing a scatterplot.

```{r}
library(tidyverse)
ggplot(mpg, aes(displ, hwy)) +
  geom_point()
```

Want to fit a "best line" of the form
$$
hwy = \beta_0 + \beta_1 disp
$$

1.  The best line is defined by a least-squares procedure -- want to obtain the intercept and slope of this best line.

2.  Suppose the best line coefficients are ($b_0, b_1)$.  Then given a value of ```disp``` we can predict a car's mileage (these are called fitted values or predicted values):
$$
\hat{hwy} = b_0 + b_1 disp
$$

3.  The residual is the difference between the actual mileage and the predicted mileage:
$$
r = hwy - \hat{hwy}
$$

#### The ```lm``` function

The ``lm`` function is a popular R function for fitting regression models.

The basic syntax is

lm(y ~ x, data = df)

where y is the response variable, x is the predictor, and df is the data frame containing the two variables.

Here I want to predict highway mileage (hwy) using the engine displacement (disp) and the data frame is mpg.

```{r}
lm(hwy ~ displ, data = mpg)
```

I use ggplot to plot the data and overlay the line of best fit.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_abline(intercept = 35.698, slope = -3.531)
```

If you save the lm(), you obtain a lm object that is named fit.

```{r}
fit <- lm(hwy ~ displ, data = mpg)
```

This is a special list containing many calculations in this regression.

Learn about components of a lm object by the attributes() function.

```{r}
attributes(fit)
```

Extract the coefficients of the best fitting line.
```{r}
fit$coef
```

#### Apply some generic functions to a ```lm``` function

print, summary, plot

#### Questions -- Playing with the lm function

Consider the ```faithful``` dataset -- can one predict the waiting time to the next eruption (variable ```waiting```) from the eruption time (variable ```eruptions```)?

#### Question 1

Construct a scatterplot of eruptions (horizontal) against waiting (vertical).
```{r}
library(tidyverse)

eruption = faithful$eruptions
waiting = faithful$waiting 
head(cbind(eruption, waiting))

plot(eruption, waiting, col="yellow", main="Scatterplot showing eruptions against waiting")

```

#### Question 2

Using the lm function find the equation of the best line for predicting waiting time from the eruption time.  Add the best line of fit to your scatterplot.

```{r}
plot(eruption, waiting, col="green", main="Scatterplot for predicting waiting and eruption")
abline(lm(waiting ~ eruption))
```


#### Question 3

Rerun the lm() function, storing the output to the variable fit.

```{r}
plot(eruption, waiting, col="blue", main = "Scatterplot of the variable fit")
fit <- abline(lm(waiting ~ eruption))
```


#### Question 4

Display the first five predicted values and the first five residual values.

```{r}
fit <- lm(waiting ~ eruption)

head(fit$fitted.values)
head(fit$residuals)
```

#### Question 5

Use the summary function on the variable fit to get some new regression output.  Explain what additional information is displayed.

```{r}
fit <- lm(waiting ~ eruption)
summary(fit)

```
The summary above shows the detailed information of the performance and coefficients. The value of the intercept eruption is: a + (b*m) ==> 28.98 + (34.09*m), this makes the "m" constant. The summary above also shows that the linear model is a good fit because the Multiple R-squared is 0.8115 which is near 1. 



