---
title: "Assignment 3a"
author: "Olaleye Damilare"
date: "9/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse)
```

From the book (Chapter 1) ...

#### Aesthetics

1. What’s gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) +
geom_point(
mapping = aes(x = displ, y = hwy, color = "blue")
)
```

Answer: The points are not blue because color would have to be put outside of aethetics. When color is put inside mapping, the value for it must be a variable. For instance, color = year would show the relationship of year with hwy and displ.
```{r}
ggplot(data = mpg) +
geom_point(color = "blue", mapping = aes(x = displ, y = hwy, color = "blue"))

```


2. Which variables in mpg are categorical? Which variables are
continuous? (Hint: type ?mpg to read the documentation for the
dataset.) How can you see this information when you run mpg?

Answer: Categorical variables : model , trans , drv, fl , class
        Continuous variables  : displ, cty, hwy, cyl, year


3. Map a continuous variable to color, size, and shape. How do
these aesthetics behave differently for categorical versus continuous
variables?

```{r}
library(ggplot2)
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy, color = year, size = drv, shape = trans))
```
Continuous variables cannot be assigned to shape. 
Categorical variables can be and R uses different shapes/colors/sizes for the different categories

4. What happens if you map the same variable to multiple aesthetics?

```{r}
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy, color = drv, size = drv, shape = drv))
```
When the same variable is assigned to multiple aesthetics, R displays all the characteristics of the aesthetics with respect to the variable.


5. What does the stroke aesthetic do? What shapes does it work
with? (Hint: use ?geom_point.)

Answer: Stroke adds an outline to the points or dots in the geom_point and geom_count. 
        The value of stroke determines the thickness of the outline.

#### Facets

1. What happens if you facet on a continuous variable?

Answer: A continous variable creates multiple facets for all the unique values.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(. ~ cty)
```
A continous variable creates multiple facets for all the unique values.


2. What do the empty cells in a plot with facet_grid(drv ~ cyl)
mean? How do they relate to this plot?

```{r}
ggplot(data = mpg) +
   geom_point(mapping = aes(x = drv, y = cyl))
```
Answer: It means that the empty cells are present as there is no value that represents the relationship between drv and cyl.

3. What plots does the following code make? What does . do?

```{r}
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy)) +
facet_grid(drv ~ .)
```
Answer: 
```{r}

ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy)) +
facet_grid(. ~ cyl)

```


```{r}
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy)) +
facet_grid(. ~ cyl)
```

4. Take the first faceted plot in this section:

```{r}
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy)) +
facet_wrap(~ class, nrow = 2)
```

What are the advantages to using faceting instead of the color
aesthetic? What are the disadvantages? How might the balance
change if you had a larger dataset?

Answer: It is difficult to resolve more than a dozen or so discrete colours, but we can have a larger number of facets than that. On the other hand, facets can be harder to read at a glance, or if the cells being compared aren't lined up in the required dimension. So in a situation like this, colours are probably better, but if we had more classes, or wanted to use colour for a different variable, facets would come into their own.
