---
title: "2 ACTION"
author: "Damilare Olaleye"
date: "9/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

Answer: 

Line chart : geom_path()/geom_line()
Boxplot    : geom_boxplot()
Histogram  ; geom_histogram()
Area Chart : geom_area()



2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions:

```{r}
library(tidyverse)
ggplot(
data = mpg,
mapping = aes(x = displ, y = hwy, color = drv)
) +
geom_point() +
geom_smooth(se = FALSE)
```

Answer: The code would produce a scatterplot with a rough line that goes through the points. The x axis is displ and y is hwy, and the points and the lines are of different colors based on the categories of drv.


3. What does show.legend = FALSE do? What happens if you
remove it? Why do you think I used it earlier in the chapter?

Answer: Legend specifies what the elements in a graph represents. For instance, different colors in a line graph could represent different entities and the names/labels of those are shown on the side of the graph as graph legend. So, setting show.legend = FALSE would hide this information.
Using legend is helpful as it makes the elements in a graph more informational. Legend explains the mapping between levels and values.


4. What does the se argument to geom_smooth() do?

Answer: According to the documentation, se stands for smooth edges. se adds error bands to the lines.

5. Will these two graphs look different? Why/why not?

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
geom_point() +
geom_smooth()

ggplot() +
geom_point(
data = mpg,
mapping = aes(x = displ, y = hwy)
) +
geom_smooth(
data = mpg,
mapping = aes(x = displ, y = hwy)
)

Answer: The two graphs will look the same. 

On the the first ggplot execution, the arguments inside of ggplot uses the data mpg, maps x and y to displ and hwy respectively. The addition of geom_point() and geom_smooth specifies which graphs to use in the execution.

The second ggplot execution specifies the arguments and the mapping inside of geom_point and geom_smooth which results in the same graph as the one above as all the mapping and the data used is same as the first one.


6. Re-create the R code necessary to generate the following graphs.

Answer:

1.
```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point() + geom_smooth(se = FALSE)
```

2.
```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point() + geom_smooth(se = FALSE, mapping = aes(group = drv))

```

3. 
```{r}
ggplot(data = mpg, aes(x = displ, y = hwy, color = drv)) + geom_point() + geom_smooth(se = FALSE)

```

4.
```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(mapping = aes(color = drv)) + geom_smooth(se = FALSE)

```

5.
```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(mapping = aes(color = drv)) + geom_smooth(se = FALSE, aes(linetype = drv))

```

6.
```{r}
  ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(color = "white", size = 4) + geom_point(aes(color = drv))
```

















