---
title: "Chapter_3_Matrices_Assignment"
author: "Damilare Olaleye"
date: "02/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Defining a Matrix

#### Question 1

The wikipedia entry https://en.wikipedia.org/wiki/Magic_square describes a magic square where the sum of integers along each row, along each column, and the diagonal is the same number.  
This page shows a 3 by 3 magic square -- put the values of this square in a matrix A with 3 rows and 3 columns.

```{r}
M <-c(2,7,6,9,5,1,4,3,8)
A <- matrix(M,nrow= 3, ncol = 3, byrow = T) 
print(A)
```

Display the 2nd row of your matrix, and the 3rd column of the matrix.

```{r}
A[2, ,drop=F]
A[, 3,drop=F]
```

#### Matrix Operations

#### Question 2

Multiply each entry of your matrix A by 3 and add 4 to the result.

```{r}
(A * 3) + 4
```

What happens if you type A * A?
It multiply itself through an entry wise multiplication 

```{r}
A * A
```

Suppose, the matrix B is defined as follows.  Can you add matrices A and B?  If not, can you redefine the matrix B so that you can add A and B?

```{r}
B <- matrix(1:12, 3, 4)
```
No, we cannot add matrices A and B, but after redefing matrix B we should be able to add A and B, as shown below: 
```{r}
A + B[1,3]
```

#### Multiplying matrices

#### Question 3

Below I have defined three matrices, m1, m2, and m3.

```{r}
m1 <- matrix(1:4, 2, 2)
m2 <- matrix(c(2, 1, 3, 5, 2, 1), 2, 3)
m3 <- matrix(1:3, 3, 1)
```


Without doing any R work, can you do the following matrix operations?

m1 %*% m2 = yes... using linear algebra logic M1 = 2 X 2, and M2 = 2 X 3... therefore, we since the 2nd col of m1 match the first col of row 1, the operations should work

m2 %*% m1 =  no... using linear algebra logic M1 = 2 X 2, and M2 = 2 X 3... therefore, we since the 2nd col of m2 does not match the first col of m1, the operations will not work. 

m3 %*% m2 =  No... using linear algebra logic M3 = 1 X 3, and M2 = 2 X 3... therefore, we since the 2nd col of m3 does not match the first col of m2, the operations will not work

m2 %*% m3 =  yes... using linear algebra logic M3 = 3 X 1, and M2 = 2 X 3... therefore, we since the 2nd col of m2  match the first col of m3, the operations will work


Check to see if your answers are correct.

```{r}
m1 %*% m2
```
```{r}
#m2 %*% m1
```

```{r}
#m3 %*% m2
```

```{r}
m2 %*% m3
```


#### Filtering on Matrices

#### Question 4

Explain in words what these filtering operations mean.

```{r}
A[A > 3]
```
To check if A is greater than 3, display numbers greater than 3.

```
A[A[1, ] > 3, ]
```
this extract the second and third row of the matrices
```
A[ , A[, 2] == 5]
```

this extract the second row of the matrices, while it checks if row 2 is equal to 5, display row 2.

#### Using apply on Marathon data

Here you are reading in some 2011 Boston Marathon data.

```{r}
library(tidyverse)
boston <- read_csv("http://www-math.bgsu.edu/~albert/3440/marathon2011.csv")
```

Extract the split times from the data frame (remove variables Age and Gender1F2M):

```{r}
split <- as.matrix(boston[, -c(1:2)])
```

#### Question 5

Using the apply function compute the mean time of each split (K0.5, K5.10, etc).  Graph these means.

```{r}
k<-apply(split, 2, mean)
k
hist(k, col="Green", main = "Compute the mean time of each split")
```


Using the apply function compute the max time of each split (K0.5, K5.10, etc).  Graph these means.
```{r}
k<-apply(split, 2, (max))
k
hist(k, col="blue", main = "Compute the max time of each split")
```


Find the index of the max split time for each runner.  Graph these index files?

```{r}
k<-apply(split, 1, which.max)
k
hist(k, col="yellow", main = "Compute the max split time for each runner")
```


What if you apply with a function that has two outputs?  To check, the range function has two outputs (the min and max).  Apply the range function for each split.

```{r}
apply(split, 2, function(x) range(min(x), max(x)))
```


