---
title: "8 Many Regressions"
output: html_document
author: "Damilare Olaleye"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Gapminder Regression Assignment

1. The gapminder package contains interesting data about many countries for many years. Install the package and load it into R.

```{r}
library(gapminder)
gapminder

aggregate(lifeExp ~ continent, gapminder, median)
```

We'll focus on the following variables from the gapminder data frame:

country = Country Name
year = Year
lifeExp = life expectancy
We're interested in exploring how life expectancies have changed over time, and how these changes vary between countries.

2. Write a short function myfunc() that will input the country name and output the mean life expectancy (over years).

Here's an outline of what you do when you write this function.

- find a subset of gapminder using the rows corresponding to the specific country
- find the mean of lifeExp for this subset

To check your work, if you type

myfunc("Albania")

you should get 68.43292

```{r}
myfunc = function(find_mean){
  round(mean(gapminder$lifeExp[gapminder$country==find_mean]), digits = 5)
}

myfunc("Albania")
```

3. Use the lapply function to find the mean life expectancy for all countries.
- First you need to create a vector containing all of the country names. (You only want to include each country a single time in your vector.)
_ Then you lapply with arguments this vector and myfunc.
```{r}
find_mean = unique(gapminder$country)
lapply(gapminder$country[find_mean], myfunc)
```



4. Instead of using lapply, use the sapply function -- how is this an improvement over using the lapply function?

```{r}
find_mean2 = unique(gapminder$country)
sapply(gapminder$country[find_mean2], myfunc)
```
Sapply displays result in columns while Lapply displays result in rows. It is also much more flexible to see the display on Sapply compared to Lapply becuase I scrolled down to see the last number on lapply while I didnt have to scroll down to see result on sapply. 


5. Now we are looking at the change of life expectancy over years. This function will do this for all countries and all years. I am fitting the regression model

lifeExp = beta0 + beta1 * (year - 1952)

When I fit this model,

beta0 is the estimated lifeExp in 1952
beta1 is the change in lifeExp for each additional year.
lm(lifeExp ~ I(year - 1952), data = gapminder)

Run this function and interpret the coefficients that are displayed.

```{r}
lm(lifeExp ~ I(year - 1952), data = gapminder)
```
Answer: The (Intercept) coefficients represent the life expectancy during the year 1952 (50.5121), while the I(year - 1952) represents the life expectancy after the year 1952 (0.3259).

6. Write a new function called onereg() that will input a country name and the output will be the coefficients of a regression fit.

If you type, for example,

onereg("Albania")

you should get

(Intercept) I(year - 1952)
59.2291282 0.3346832

```{r}

onereg <- function(d){
lm(lifeExp ~ I(year - 1952), data = d)
}
by(gapminder, gapminder$country, onereg)
```


7. Use sapply with your list of countries and the onereg function to fit regression models for all countries. Save your output in a variable manyfits.

```{r}
#Error in eval(predvars, data, env) : object 'lifeExp' not found
#not sure about why I'm getting error here, but I'm very sure that I'm right

#manyfits<-sapply(find_mean, onereg)
```

8. Construct a scatterplot of the intercepts and slopes for all countries.

```{r}
plot(lm(formula = lifeExp ~ I(year - 1952), data = gapminder))
```


9. Identify the country with the largest intercept and the country with the largest slope value.


The country with the largest intercept is: Swedeen with 71.6050 and the country with the largest slope is: Vietnam with 0.6716 
