---
title: "Assignment 4b"
output: html_document
author: "Olaleye Damilare"
date: "9/20/2019"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in the readr and ggplot2 packages.

```{r}
library(readr)
library(ggplot2)
library(dplyr)
```



```{r}
earthquakes <- read_csv("earthquakes.csv")
```


1.) Create a new data frame of only the earthquakes below the equator.

```{r}
below.earthquake <- earthquakes %>% filter(location.latitude < 0)
below.earthquake
```

2.) Create a new data frame of only the earthquakes above the equator.

```{r}
above.earthquake <- earthquakes %>% filter( location.latitude > 0)
above.earthquake
```

3.) Use the arrange verb to find the most severe earthquake

```{r}
most.severe <- earthquakes %>% arrange(desc(impact.magnitude))
head(most.severe,1)
```

4.) Are there more earthquakes above or below the equator?
```{r}
above <- count(above.earthquake)
above
```
```{r}
below <- count(below.earthquake)
below
```
```{r}
count <- ifelse (above > below, "There are more Earthquakes above", "More Earthquakes below")
count
```

5.) Find all of the earthquakes in June – where was the most severe one.

```{r}
earthquakes %>% filter( time.month == 6 )
```
There are no earthquakes in June



6.) Find all of the earthquakes in December 
```{r}
decdata <- earthquakes %>% filter( time.month == 12 )
decdata
```
There are no earthquakes in December 

6b.) – where was the most severe one.
```{r}
severe<-decdata %>% arrange(desc(impact.magnitude))
head(severe,1)
```
7.) (Extra credit) How many earthquakes listed are in Ohio?
```{r}
location <- earthquakes %>% filter( location.name == "Ohio" )
location
```

