---
title: "Assignment 6a"
author: "Damilare Olaleye"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The Lahman database is a collection of season-to-season statistics for all baseball players in MLB history.  It is available in the ```Lahman``` package.  Before you start this assignment install the ```Lahman``` package.

To start, load in the ```Lahman``` and ```tidyverse           ``` packages.

```{r}
library("tidyverse")
library("Lahman")
```

#### Creating two top-ten lists.

#### Top-ten batting averages.

**Choose a year when one of your parents was born -- what year did you choose?**
1977

**Use filter to choose batting data for the season you chose.**

```{r}
battingData <- data.frame(Batting)
battingData <- filter(battingData, yearID == "1977")
```

**Grouping data by playerID, find the total number of AB and H for each player.**

```{r}
battingData <- battingData %>% group_by(playerID) %>% summarise(AB = sum(AB),H = sum(H))
```

**Use mutate to compute the batting average, rounding to 3 decimal places.**

```{r}
totalAB_H <- battingData %>% mutate(battingAvg = round(H / AB, 3))
```

**Use filter to limit list to qualifying players with 400 AB.**

```{r}
bat400 <- totalAB_H %>% filter(AB > 400)
```

**Arrange by decreasing batting average.**

```{r}
bat400 <- bat400 %>% arrange(desc(battingAvg))
```

**Use slice to only show top 10 players.**

```{r}
slice(bat400, 1:10)
```

**By use of the inner_join function, merge your data with the first and last names of the players found in the Master data frame.**

```{r}
batt_10 <- Master %>% select(playerID, nameFirst, nameLast) %>% inner_join(bat400, Master, by = "playerID")
```


**Display your list to compare with the list on Baseball-Reference.**

```{r}
batt_10 <- batt_10 %>% arrange(desc(battingAvg)) %>% select(nameFirst, nameLast, BA = battingAvg)
slice(batt_10, 1:10)
```

#### Top 10 WHIP Values

WHIP stands for the mean number of walks and hits allowed per inning for a pitcher.  In the Pitching data frame, BB is the number of walks, H is the number of hits, and IPouts is the number of outs.  Then WHIP is defined by

$$
WHIP = \frac{BB + H}{IPouts / 3}
$$

For example, in 2016, Clayton Kershaw allowed 97 Hits and 11 Walks and generated 447 outs and his WHIP was 

$$
WHIP = \frac{11 + 97}{447/3} = 0.725
$$

**Find and display the top 10 WHIP pitchers for your season** (qualifying pitchers are the ones where IPouts >= 486).
```{r}
WHIP_data <- data.frame(Pitching)
WHIP_data <- WHIP_data %>% filter(yearID == 1977)
WHIP_data <- WHIP_data %>% group_by(playerID) %>% 
                           summarise(BB = sum(BB), H = sum(H), IPouts = sum(IPouts))
WHIP_top <- WHIP_data %>% mutate(WHIP = round((BB + H)/(IPouts/3), 3)) %>% 
                          filter(IPouts >= 486)
WHIP_top <- Master %>%
            select(playerID, nameFirst, nameLast) %>% 
            inner_join(WHIP_top, Master, by = "playerID") %>% 
            select(nameFirst, nameLast, WHIP)
WHIP_top <- WHIP_top %>% arrange(WHIP)
slice(WHIP_top, 1:10)
```

