---
title: "Assignment 10b"
author: "Damilare Olaleye"
date: "November 1st, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
library(tidyverse)
library(htmlwidgets)
```
Load file
```{r}
license_data <- read_csv("http://www-math.bgsu.edu/~albert/text/files/licenseplates.csv")
plates <- license_data$Plate
```

1.  Find all license plates that begin with B.
```{r}
str_view_all(plates, "^B", match = T)

```

2.  Find license plates that end with 4.
```{r}
str_view_all(plates, "4$", match = T)

```

3.  Find license plates with 2 consecutive B's.
```{r}
str_view(plates, "BB", match = T)
```

4.  Find license plates with between 3 and 4 consecutive B's.
```{r}
match_B <- c("BBB", "BBBB")
str_view_all(plates, match_B, match = T)
```


5.  Find all license plates that are of the form  "num, letter, num, letter, num, letter".
```{r}
str_view_all(plates, "\\d[^\\d]\\d[^\\d]\\d[^\\d]", match = T)
```

6.  Find all license plates that start with a number and end with a letter.
```{r}
str_view_all(plates, "\\d....[^\\d]", match = T)
```

7.  Find all license plates that either start or end with "AB".
```{r}
str_view_all(plates, "^AB|AB$", match = T)
```

8.  Find all license plates that contain one or more (AB)s.
```{r}
str_view_all(plates, "AB", match = T)
```

9.  Find all license plates that contain the pattern "a number, a number".
```{r}
str_view_all(plates, "\\d\\d", match = T)
```

10.  Find the license plates that are all letters.
```{r}
str_view_all(plates, "[^\\d][^\\d][^\\d][^\\d][^\\d][^\\d]", match = T)
```
11. Find the license plates that contain AA, BB, or CC.
```{r}
str_view_all(plates, "AA|BB|CC", match = T)
```

12. Find the license plates that contain only one letter.
```{r}
str_view_all(plates, "[^\\d][\\d][\\d][\\d][\\d][\\d]|[\\d][^\\d][\\d][\\d][\\d][\\d]|[\\d][\\d][^\\d][\\d][\\d][\\d]|[\\d][\\d][\\d][^\\d][\\d][\\d]|[\\d][\\d][\\d][\\d][^\\d][\\d]|[\\d][\\d][\\d][\\d][\\d][^\\d]", match = T)
```

