---
title: "Assignment 5a"
author: "Damilare Olaleye"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Data Manipulation with dplyr

We'll work with a dataset ```flights``` on all 336,776 flights departing from New York in 2014

**Get this data by installing and loading the ```nycflights13``` package.**

```{r}
## install and load the nycflights13 package
library(nycflights13)
```

Your instructor will ask you to look at a specific row of the flights data.  

**What was the carrier for this flight?  What day was this flight?  When was it scheduled to depart?  Did it depart on time?** 

```{r}
## save row of data frame flights to variable FLIGHT
## display FLIGHT to answer questions

flights[99, ]
```

**Load the tidyverse package that includes the dplyr package.**

```{r}
## load tidyverse
library(tidyverse)
```

Basic dplyr verbs:  filter, arrange, select, mutate, summarize

The ```filter``` verb

**Describe briefly what filter does in a data frame**

**Use filter to find the flights that departed on May 5**

```{r}
# flights departing on May 5
filter(flights, month == 6, day == 5)
```

**Save the "May 5" flights data frame in the data frame may5 **

```{r}
## save may 5 flights in data frame called may5
may5 <- filter(flights, month == 6, day ==5)
may5
```

**Find the flights where the actual departure time is after 9 pm**

```{r}
## flights with actual departure after 9 pm
filter(flights, dep_time > 2100)

```

**Find the flights that depart between 10 and 20 minutes late**

```{r}
## flights that depart between 10 and 20 minutes
filter(flights, (dep_time - sched_dep_time >= 10 & dep_time - sched_dep_time <= 20))
```

**Find the flights during February or April**

```{r}
## flights during February or April
filter(flights, (month == 2 | month == 4))

```

**Find the flights that are not Delta**

```{r}
## flights that are not Delta
filter(flights, carrier != "DL")
```

**Find the flights that arrived more than two hours late, but didn't leave late**

```{r}
## flights arriving two hours late but didn't leave late
filter(flights, dep_delay <= 0, arr_delay > 120)
```

**Find the flights that were delayed at least an hour, but made up over 30 minutes in flight**

```{r}
## flights delayed at least an hour but made up over 30 minutes in flight
filter(flights, dep_delay >= 60 & arr_delay <= 30 )
```




