---
title: "Assignment on Factors"
author: "Damilare Olaleye"
date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(LearnBayes)
```

Assignment on Factors

The dataset studentdata in the LearnBayes package contains measurements from 657 students in an introductory statistics class.
```{r}

head(studentdata)
```


1. Examine the factor variable Drink. Find the values, levels of this variable.

```{r}
drink_variable = data.frame(studentdata)
str(drink_variable['Drink'])
```


2. Try replacing the first value of Drink to "tea". You should get an error?  Fix this so you don't get an error?

```{r}
#This displays an error
#drink_variable[1] = "tea"
#drink_variable

#drink_variable$m.Drink = with(drink_variable,drink_variable="tea",stringsAsFactros=F)

drink_variable$Drink = (factor(drink_variable$Drink, levels = c("milk","pop","water","tea")))
levels(as.factor(drink_variable$Drink))

drink_variable$Drink[1] = 'tea'
head(drink_variable['Drink'])
```


3. Using the tapply function, find the median number of pairs of shoes owned by men and women.

```{r}
using_tapply <- with(drink_variable[complete.cases(drink_variable),], tapply(Shoes, Gender, median))
using_tapply
```


4. Using the tapply function, find the median number of pairs of shoes owned by students classified by gender and favorite drink.

```{r}
classfier_tapply <- with(drink_variable[complete.cases(drink_variable),], tapply(Shoes, list(Drink, Gender), median))
classfier_tapply
```


5. Using the split function, split the student heights in two groups by gender -- store the output in the variable Heights.

```{r}
Heights <- with(drink_variable[complete.cases(drink_variable), ], split(Height, Gender))
head(Heights)
```

6. Using the lapply function, find the maximum height for each gender. (By the way, there is some missing values, so you will have to remove them by the na.rm = TRUE option.) Are you surprised by the answer? Why?

```{r}
lapply(Heights, max)
```


7. Construct a bar graph of the Drink variable.

```{r}
barplot(main="Drink Variable Bar Graph",cex.main=1,col="blue",col.main="blue",table(drink_variable$Drink))
```


8. By changing the order of the levels of the factor variable, redraw the bar graph so that the frequencies are decreasing from left to right.

```{r}
drink_variable$Drink = (factor(drink_variable$Drink, levels = c("tea","milk","pop","water")))
barplot(main="Freq increasing from Left to Right",cex.main=1,col.main="green",col= "green",table(drink_variable$Drink))
```


9. Use the table function to tabulate all of the students by Gender and Drink choice.

```{r}
table(list(drink_variable$Gender, drink_variable$Drink))
```


10. By using the prop.table() function, find the proportion of men who prefer milk and the proportion of women who prefer milk.

```{r}
prop.table(table(list(drink_variable$Gender, drink_variable$Drink))[,4])
```


11. Define a new variable Sleep in the studentdata data frame that is equal to the difference in the WakeUp and ToSleep times.

```{r}
drink_variable['Sleep'] = drink_variable$WakeUp - drink_variable$ToSleep
drink_variable$Sleep
```


12. What is the mean amount of sleep of these students?
```{r}
mean_amt <- mean(drink_variable$Sleep, na.rm = T)
mean_amt
```


13. Use the cut function to classify students as Light or Heavy sleepers if they get less than or more than 7.5 hours of sleep respectively -- call the new variable CSleep.

```{r}
drink_variable['CSleep'] <- cut(drink_variable$Sleep, breaks = c(-Inf,7.5, Inf), labels = c('Light','Heavy'))
```


14. Use the table function to classify all students by Gender, Drink choice, and CSleep.

```{r}
classfiy_students <- table(list(drink_variable$Gender, drink_variable$Drink, drink_variable$CSleep))
classfiy_students
```


15. Graph the counts in the table constructed in 14.
```{r}
plot(main="Classify all students by Gender, Drink choice, and CSleep",col="yellow",classfiy_students)
```


16. Using the apply function three times, find the three marginal tables of Gender by Drink, Gender by CSleep,and Drink by CSleep.
```{r}
apply(table(drink_variable$Gender, drink_variable$Drink), 1, sum)
apply(table(drink_variable$Gender, drink_variable$CSleep), 1, sum)
apply(table(drink_variable$Drink, drink_variable$Sleep), 1, sum)
```


17. By using code, identify the combination of (Gender, Drink, CSleep) that has the highest frequency.
```{r}
highest_freq<-data.frame(table(drink_variable$Gender, drink_variable$Drink, drink_variable$CSleep))
 highest_freq[highest_freq$Freq == max(highest_freq$Freq), ]
 
```


 