---
title: "Baseball_Cards, Assignment 1b"
author: "Jim Albert, Damilare Olaleye"
date: "August 21, 2017, August 29, 2019 "
output: html_document
---

## Baseball Card Collection Problem

Your instructor used to collect baseball cards which motivates this simulation exercise.

Suppose a set of baseball cards are numbered from 1 to N.  You purchase M cards where each card is equally likely to be numbered 1 through N.

What is the chance you buy a complete set?

Set values of N and M:

```{r}
N <- 30
M <- 150
```

We buy M cards, where each card is equally likely to be 1, ..., N.  Store the purchases in the vector purchased.

```{r}
purchased <- sample(N, size=M, replace=TRUE)
purchased
```

Graph the numbers of the purchased cards.

```{r}
library(ggplot2)
ggplot(data.frame(Card_Number=purchased), 
       aes(Card_Number)) +
  geom_bar()
```

We identify the unique cards that are purchased.  If the number of unique cards is equal to N, we have a complete set; otherwise the set is not complete.

```{r}
U <- unique(purchased)
ifelse(length(U) == N, "complete set", "not complete")
```

I repeat this exercise 100 times -- each time, I purchase N cards, identify the unique ones, and see if I have a complete set.  When I am done, I have a vector of "complete set" and "not complete".

```{r}
R <- replicate(100, expr = {
  purchased <- sample(N, size=M, replace=TRUE)
  U <- unique(purchased)
  ifelse(length(U) == N, "complete set", "not complete")
})
R
```

I construct a frequency table of the outcomes and graph the data.

```{r}
(freq <- table(R))
ggplot(data.frame(R), aes(R)) + geom_bar()
```

#### Conclusion

With the values of N and M chosen, it is pretty likely that I have a complete set.