---
title: "7 Love Songs"
output: html_document
author: "Damilare Olaleye"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(stringr)
```


```{r}
require(ggplot2)
dat<- data.frame(t=seq(0, 2*pi, by=0.1) )
 xhrt <- function(t) 16*sin(t)^3
 yhrt <- function(t) 13*cos(t)-5*cos(2*t)-2*cos(3*t)-cos(4*t)
 dat$y=yhrt(dat$t)
 dat$x=xhrt(dat$t)
ggplot(dat, aes(x, y)) + theme_minimal() +
geom_polygon(fill = "red") +
annotate(geom = "text",x = 0, y = 0,
label = "Happy Valentine's \nDay!",
size = 12, color = "white")
```

1. Find the lyrics to two of your favorite love songs. Store the lyrics to the two songs in separate text files in your current working folder.

```{r}
song_one <- read.delim("loveSongs1.txt")
song_one

song_two <- read.delim("loveSongs2.txt")
song_two
```


2. Use the findwords( ) function to find the locations of all of the words in your first love song -- store these locations in the variable song1.

```{r}
findwords <- function(loveSong1) {
  txt <- scan(loveSong1, "")  
  txt <- tolower(gsub('[[:punct:]]+','', txt))
    wl <- list()  
    for (i in 1:length(txt)) {
      wrd <- txt[i]
      wl[[wrd]] <- c(wl[[wrd]], i)  
    }   
    return(wl)
}

findwords("loveSongs1.txt") -> song1
head(song1)

```

3.  Find the locations of the word "love" in your song.

```{r}
song1$love

```

4.  By use of the lapply function, find the number of times each word is used in your song.

```{r}
lapply(song1, length)
```

5.  By use of the lapply function, find the location of the first time each word is used in your song.
```{r}

lapply(song1, function(elem) elem[1])

```


6.  By use of the function alphawl(), sort your list alphanumerically.  Demonstrate that it works.
```{r}
alphawl <- function(wrdlst){
 nms <- names(wrdlst) 
 sn <- sort(nms)
 return(wrdlst[sn])
}

alphawl(song1)
```


7.  By use of the function freqwl(), sort your list by the word frequency.  Demonstrate that is works.
```{r}
freqwl <- function(wrdlst){
  freqs <- sapply(wrdlst,length)
  return(wrdlst[order(freqs)])
}

freqwl(song1)
```


8.  Repeat questions 2 and 3 for your second song. Compare the love in your two songs.
```{r}
# question 2

findwords <- function(loveSong2) {
  txt <- scan(loveSong2, "")  
  txt <- tolower(gsub('[[:punct:]]+','', txt))
    wl <- list()  
    for (i in 1:length(txt)) {
      wrd <- txt[i]
      wl[[wrd]] <- c(wl[[wrd]], i)  
    }   
    return(wl)
}

findwords("loveSongs2.txt") -> song2
head(song2)



```

question 3
```{r}
song2$love
```

Comparing the love

love appeared 8 times in song 2 compared to song 1 where love appeared 15 times


