---
title: "classSpt17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(ggplot2)
```

```{r}
opiod <- read_cvs("graduates.py")
```


Lets focus on a particular major and see how the number of graduates with taht major has changed over years.

Which major do you want to look at? (class response) -> Computer Science
It appears that is coded as "Computer Science and Math."

Create a new dataset cs:

```{r}

cs <- grad[grad$Education.Major == "Computer Science and Math", ]
```

Graph the number of majors over time. 

```{r}
ggplot(cs, aes(Year, Demographics.Total)) + geom_point()
```

I see ...

-Number of cs majors seem to be increasing over time. 
-More rapid incerease in the number of cs majors in later years.
-What is going on in 2002,2006, and 2008?

MIght be intresting to look at other majors to see if there is a smiar pattern of weird data. 

Can we see how the gender distribution of cs majors has changed over years?

Compute a new variable   which gives the fraction of femlae majors.

```{r}
cs$Fraction_Females <= cs$Demographics.Gender.Females /cs$Demographics.Gender.Females + cs$Demographics.Gender.Males)

```

Graph these fractions against year.
```{r}
ggplot(cs, aes(Year, Fraction_females)) + geom_point()
```

Are there any categorical variables of interest?

Look at the median salaray of all these majors for a spceific year.

```{r}
d2015 <- grad[grad$Year == 2015,]
```

Graph the median salaries.
```{r}
ggplot(d2015, aes(Education.Major, salaries.Median)) + geom_point() + coord_flip()
```


