---
title: "Vectors and Data Frames"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Introduce data types

Different data types such as numeric, character, logical, etc.

Data for the day is a table of the top grossing movies for 2019. 

#### Vector 

```{r}
title <- c("Avengers", "The Lion King", "Toy Story 4", "Captain Marvel", "Spider-Man", "Aladdin", "Us", "John Wick", "How to Train Your Dragon", "Secret Life of Pets 2")


```

Display this vector: 
```{r}
title
```

Define a numeric vector -- total gross in millions,

```{r}
gross <-c(858, 512, 427, 426, 380, 353, 175, 170, 161, 157)
```

Working with vectors:

use square brackets to index specific values.

```{r}
1:3
title[1:3]
gross[c(2, 5, 7)]
```

Illustrate some functions on vectors
```{r}
mean(gross)
median(gross)
max(gross)
sd(gross)
```

These functions only work for numeric data
```{r}
mean(title)
median(title)
```

```{r}
sqrt(-2)
```

Try sorting ...
```{r}
sort(title)
```

If you sort alphabetically, what is the 4th movie title?

```{r}
sort(title)[4]
```

Introduce Logical Variables.

Which movies gross more than 300 million?

Identify the movies that satisfy this condition 

```{r}
large <- gross > 300
```

Take title variable -- subset using the logical vector. 

```{r}
title[large]
```

Some operations with vectors.

Write down the gross revenues in dollars. 

```{r}
newgross <- gross * 1e6
newgross
```

Suppose each of the top ten movies in 2020 earn 10 more million dollars

```{r}
gross + 10
```

work with two numeric vectors

```{r}
opening <- c(357, 191, 120, 153, 92, 91, 56, 55, 46)
```

What fraction of the total gross was earned in the opening weekend? Which movie had the highest fraction?

```{r}
frac <- opening / gross 
title[frac == max(frac)]
```

Some times you want to work with vectors of different sizes

```{r}
x <- 1:4
y <- 1:5
```

```{r}
x+y
```

```{r}
x
y
```

Data Frame

A data frame is a rectangular array of data where rows are cases (people) and the colums are variables which can be different types. 

```{r}
(movies <- data.frame(title, opening, newgross))
```

selecting portions of data frame. 
use the bracket notation.

```{r}
movies[2, 3]
```

```{r}
movies[2,]
```

```{r}
movies[c(2, 3, 5), ]
```

```{r}
movies[c(1,5), c(2:3)]
```


Many ways to refer to varaibles in a data frame. 

```{r}
movies[, 3]
```

```{r}
movies[, "newgross"]
```

```{r}
movies$newgross
```

Try some subsetting using logical varaibles.

Choose the rows of the movies which opened for less than 100 million dollars

```{r}
movies[movies$opening < 100, ]
```
 movies[c(T,T,T, F, F, F, F, F, F, F)]
