---
title: "Final Project"
author: "Ashkan Dastani & Damilare Olaleye"
date: "12/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is data from Pronto, which is Seattle's cycle sharing system with 500 bikes and 50 stations located throughout the city.
This is data on all 2015 trips.

Some notes about the data are trips of duration less than one minute or greater than eight hours were removed from original table and

Gender and birth year are only available for Annual Members.


-starttime: day and time trip started, in PST

-stoptime: day and time trip ended, in PST

-tripduration: time of trip in seconds 

-from_station_name: name of station where trip originated

-to_station_name: name of station where trip terminated 

-usertype: "Short-Term Pass Holder" is a rider who purchased a 24-Hour or 3-Day Pass; "Annual Member" is a rider who purchased an Annual Membership

-gender: gender of rider 

-birthyear: birth year of rider

-bikeid: ID attached to each bike

-trip_id: numeric ID of bike triptrip taken

-from_station_id: ID of station where trip originated

-to_station_id: ID of station where trip terminated


## Questions

1.  How long are typical bike trips?

2.  What ages and gender are taking bike trips?

3.  What are popular stations to depart and arrive?

4.  What days and times are popular for taking bike trips?


## R Markdown

```{r}
library(tidyverse)
ds <- read_csv("http://personal.bgsu.edu/~dashkan/trip.csv")
ds %>% mutate(age = 2015 - birthyear) ->ds
ds %>% filter(trip_id != 59000) -> ds
ds
```

```{r}
ggplot(ds, aes(age)) + geom_bar(width = 1, color = "brown", fill = "orange") + ggtitle("Age of Members") + xlab("Age")
```


From the graph we can see that the count of most users are between the ages of 20 and 40. We can also see that there is a large spike just before the 30 year old mark. There are very few members in the higher age ranges. And finally, there are also almost no under 20 year old users.


```{r}
ds %>% mutate(tripinmin = tripduration/60) -> ds2
ggplot(ds2, aes(tripinmin)) + geom_histogram(color = "brown", fill = "orange", binwidth = 5) + xlab("Trip Duration in Minutes") + ggtitle("Trip Duration of Users")
```


From the graph we see that most trips are very short, and after the one hour mark the number of trips begins to decline and became very low.

```{r}
ds2 %>% filter(!is.na(gender)) -> ds3
ggplot(ds3, aes(gender)) + geom_bar() + ggtitle("Gender of Members")
```


The graph shows that the vast majority of users are Male. The combination of the femail and other category do not even come close to the number of male riders.

```{r}
ggplot(ds3, aes(gender, tripinmin)) + geom_boxplot() + coord_flip() + ylim(0,100) + ggtitle("Member Trip duration by Gender")
```


The graph shows that the female and other categories on average takes longer trips than male, the graph graph shows a very large number of outliers

```{r}
ds2 %>% group_by(age) %>% 
  summarize(N = n(), avgtripinmin = mean(tripinmin)) %>%  arrange(desc(N)) -> avgtripbyage
avgtripbyage
ggplot(avgtripbyage, aes(age, avgtripinmin)) + geom_point() + ggtitle("Average trip length by age")
```


The graph shows the avearge is below 20 minutes, and that the large averages are between 71 and 79 year old group. The 84 year olds have 1 and the 71 years old have 2 users. The 79 year old group with 1 user shows that it's just one of them taking a long trip. 

```{r}
ggplot(ds2, aes(usertype, tripinmin)) + geom_boxplot() + coord_flip() + ylim(0,100) + ggtitle("Members vs Short Term pass holders trip duration")
```


The short-term pass holders on average had the longer trips compared to the members. There are large number of outliers from the groups, but the members have more outliers than the short-term pass holder.


```{r}
ds %>% group_by(from_station_name) %>% 
  summarize(N = n()) %>%  arrange(desc(N)) ->popfromstations
popfromstations
head(popfromstations)
head(popfromstations %>% arrange(N))
```

The least popular stations to depart from are Pronto shop, 8D OPS 02, Pronto shop 2, Pier 66 / Alaskan Way & Bell St, UW McCarty Hall / Whitman Ct, and Thomas St & 5th Ave N. The most popular stations to depart from are Pier 69 / Alaskan Way & Clay St, E Pine St & 16th Ave, 3rd Ave & Broad St, 2nd Ave & Pine St, Westlake Ave & 6th Ave, and E Harrison St & Broadway Ave E. 


```{r}
ds %>% group_by(to_station_name) %>% 
  summarize(N = n()) %>%  arrange(desc(N)) -> poptostations
poptostations
head(poptostations)
head(poptostations %>% arrange(N))
```

The least popular stations to arrive at are 8D OPS 02, Pronto shop 2, Pronto shop, Pier 66 / Alaskan Way & Bell St, Broadway and E Denny Way, and UW McCarty Hall / Whitman Ct.
Un-surprisingly the least and most popular stations to depart from and arrive at are very similar. The most popular stations to arrive at are 2nd Ave & Pine St, Pier 69 / Alaskan Way & Clay St, Westlake Ave & 6th Ave, 3rd Ave & Broad St, PATH / 9th Ave & Westlake Ave, and Occidental Park / Occidental Ave S & S Washington St.


```{r}
ds %>% mutate(timeofdaystarted = starttime) -> ds4
ds4 %>% mutate(timeofdaystarted = substr(timeofdaystarted, 11, 16)) -> ds4
ds4 %>% mutate(timeofdaystopped = stoptime) ->ds4
ds4 %>% mutate(timeofdaystopped = substr(timeofdaystopped, 11, 16)) -> ds4
ds4
```

```{r}
library(lubridate)
ds4 %>% mutate(timeofdaystarted = ((as.numeric(seconds(hm(timeofdaystarted))))/60)/60) -> ds5
ds5
ggplot(ds5, aes(timeofdaystarted)) + geom_histogram(color = "brown", fill = "orange") + ggtitle("Time of Day the Cycling trip Started")
```

The graph shows that people are taking trips just before 10am or just before 8pm. There are very few late at night or very early in the morning. This however correspond to the hours you would go to or leave work, a 9am-5pm or 8am-4pm work cycle.

```{r}
ds5 %>% mutate(timeofdaystopped = ((as.numeric(seconds(hm(timeofdaystopped))))/60)/60) -> ds5
ds5
ggplot(ds5, aes(timeofdaystopped)) + geom_histogram(color = "brown", fill = "orange") + ggtitle("Time of Day the Cycling trip Ended")
```

Since most trips are very short the this graph looks very similar to the time of day that trips started. For similar reasons.

```{r}
ds5 %>% mutate(dayofweek = starttime) -> ds5
ds5 %>% mutate(dayofweek = substr(dayofweek, 0, 10)) -> ds5
ds5 %>% mutate(dayofweek = wday(mdy(dayofweek), label = TRUE)) -> ds5
ds5
ds5 %>% filter(!is.na(dayofweek)) -> ds6

```

```{r}
ggplot(ds6, aes(dayofweek)) + geom_bar(color = "brown", fill = "orange") + ggtitle("Day of the week that is popular for bike trips")
```

From the graph we can see that the least popular day to use the bike sharing is on Sunday, while Saturday is the second least popular day for a trip. All of the week days are fairly similar to each other but Tuesday is the day of the week that is popular for bike trips.